---
title: "lab2"
output: html_document
---

# Labjournal Week 2

```{r, globalsettings, echo = FALSE, warning = FALSE}

#install.packages("RSiena")
#install.packages("igraph")
library(RSiena)
library(igraph)
library(tidyverse)
library(knitr)

opts_chunk$set(warning = FALSE, message = FALSE)
# tidy = TRUE (uitlijnen/spaties)
# results = 'hold' (output op het einde)
# eval = FALSE (code niet runnen, op basis van eerder werk)
```

## Load Networks s501 & s502

```{r}
s501
s502
# ?s501 
```

## Inspect Network s501

```{r}
View(s501)
class(s501)
str(s501)

s501_copy <- s501

# Wie is bevriend met wie?
# Hoeveel vrienden hebben mensen?
# Met hoeveel mensen is iemand bevriend?

# ?graph_from_adjacency_matrix()
# ?from_adjacency()
```

## Descriptives

```{r}
friends_nominated <- rowSums(s501_copy) # nr of times someone has been nominated
friends_has <- colSums(s501_copy) # nr of friends someone has answered

View(friends_nominated)
View(friends_has)

table(friends_nominated)
table(friends_has)

barplot(friends_nominated)
barplot(friends_has)

prop.table(table(friends_nominated))
prop.table(table(friends_has))

barplot(s501_copy[1,]) # for respondent one

network <- graph_from_adjacency_matrix(
  s501_copy,
  mode = c("directed"),
  weighted = NULL,
  diag = FALSE,
  add.colnames = NA,
  add.rownames = NA
)

plot(network) # graphic overview of someone's network (rough)
```

## Calculate Clustering

### *Tutorial, Chapter 4, 4.3.2.5 Clustering*

- People lump together in separate groups

- The **clustering coefficient** for v*i* is defined as

  - The observed ties between all direct neighbors of v*i* via an ingoing and/or outgoing relation

  - Divided by all possible ties between all direct neighbors of v*i*

- k*i* are the number of neighbors of v*i*

- ∑ejk are the number of edges between these neighbors

- The local clustering coefficient is CC(vi) = ∑ejk / ki∗(ki−1)

- For undirected networks, the clustering coefficient is the same as the **transitivity index**

  - The number of transitive triads divided by all possible transitive triads

  - This does *not* appply to directed networks

### Formula

$$CC(vi) = ∑ejk / ki∗(ki−1)$$

## Clustering coefficient in R, using the igraph package

- Use the transitivity() function? For more information: <https://r.igraph.org/reference/transitivity.html>

- *"Transitivity measures the probability that the adjacent vertices of a vertex are connected. This is sometimes also called the clustering coefficient."*

- I am not sure if this function fits, as k it's for undirected networks (in which relationships are mutual)

  - *"The global transitivity of an undirected graph. This is simply the ratio of the count of triangles and connected triples in the graph. In directed graphs, edge directions are ignored"*

  - Other types are possible, but they are all undirected

  - Q: on Stack Overflow someone says the global CC is the same for directed and undirected networks; how does that work? <https://stackoverflow.com/questions/68897940/how-can-i-calculate-the-cluster-coefficients-of-a-directed-graph-in-r>

```{r}

?transitivity()

# Check if s501 is an undirected or directed network

for (i in 1:nrow(s501)) {
  if (sum(s501[i,]) == sum(s501[,i])) {
    print(paste0("participant ",
                 i,
                 ": ",
                 "equal number friends mentioned & mentioned as friend"))
  } else {
      print(paste0("participant ",
                 i,
                 ": ",
                 "numbers differ"))
  }
}

# since these values differ, it's not an undirected network
# A can nominate B as friend, without being nominated back (= non symmetrical)

transitivity(network) # 0.4229075 (but what does this number say? nothing, i think)
```

## Another approach, using the DirectedClustering package

- <https://cran.r-project.org/web/packages/DirectedClustering/DirectedClustering.pdf>

```{r}
# install.packages("DirectedClustering")
library(DirectedClustering)

ClustBCG(mat = s501_copy, type = "directed") # = 0.3994831

# Gives local clustering coefficients for each node
# And the global one, averaged across these nodes

# The LocalCC & GlobalCC are for undirecte networks
# For directed networks, there're four options (Cycle, In, Out, Total)
# https://www.sciencedirect.com/science/article/pii/S1007570424002314 
# GlobaltotalCC & totalCC (local) 

# ClustF(mat = s501_copy, type = "directed") # = 0.3994831 (same value)
```

## Check with a small network

```{r}
# ?matrix()
small_matrix <- matrix(data = NA, nrow = 4, ncol = 4, byrow = TRUE)
small_matrix_no_friends <- matrix(data = 0, nrow = 4, ncol = 4, byrow = TRUE)
small_matrix_all_friends <- matrix(data = 1, nrow = 4, ncol = 4, byrow = TRUE)

#View(small_matrix)
#View(small_matrix_no_friends)
#View(small_matrix_all_friends)

ClustF(small_matrix_no_friends, type = "directed") # GlobaltotalCC = 0
ClustF(small_matrix_all_friends, type = "directed") # GlobaltotalCC = 1

# Let's try a 
small_matrix <- matrix(data = c(0,1,1,1,
                                1,0,0,0,
                                1,1,0,0,
                                0,0,1,0), nrow = 4, ncol = 4, byrow = TRUE)

rownames(small_matrix) <- c("p1", "p2", "p3", "p4")
colnames(small_matrix) <- c("p1", "p2", "p3", "p4")

print(small_matrix)

?ClustF()

ClustBCG(mat = small_matrix, type = "directed")

```

## By hand

- Sum of all the edges between the neighbors of v*i*

- Let's say participant 1

  - As participant 1 is friends with participant 2, 3 and 4, we count the ties between them = 2

  - Next, we calculate the number of neighbors of participant 1 = 3

    \$\$

    ∑ejk / ki∗(ki−1)

    2 / 3\*(3-1)

    2 / 6 = 0,333

    \$\$

- I think this is the out-degree CC?

- If we check with the ClusterBCG output for participant 1, that tracks
